---
title: "Predicting the Top-10 HR Hitters for 2021"
author: "Greg Ngim"
date: "11/20/2020"
output: html_document
---

## Overview

In this exercise, I will be predicting the **top 10 home run hitters for the 2021 season** and the **number of HRs they will hit**. I've done a bit of work on constructing an expected home run model previously that you can check out [here](https://lemursource.wordpress.com/2018/11/02/continued-building-a-predictive-home-run-model-xhr-fb/).

I leaned on [Mike Podhorzer's xHR per fly ball rate model](https://fantasy.fangraphs.com/introducing-the-new-statcast-charged-batter-xhrfb-rate/) for inspiration. After testing a [quadratic regression model](https://thedynastyguru.com/2020/04/08/jordans-research-ramblings-expected-power-based-on-fly-ball-distance-mlb-observations/) where I squared the average fly ball distance term, I found that the R-squared and RMSE results were not an improvement over a linear regression model.

## Methodology

I chose to do things the "hard way". Instead of downloading data from Fangraphs and/or Baseball Savant, used the [baseballr package](http://billpetti.github.io/baseballr/) created by Bill Petti to scrape statcast data from the past 4 seasons (2017-2020). Then, using the dplyr package within tidyverse, calculate our required stats for modeling, visualize our results with ggplot2, interact with our data using plotly, and predict our predictors using Tom Tango's basic [Marcel projection system](https://www.baseball-reference.com/about/marcels.shtml). By predicting our predictors, I mean in order to use our xHR/FB model, we'll need to calculate the predictive rate stats for next season. I'll explain this in greater detail below.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
library(baseballr)
library(tidyverse)
library(plotly)
library(caTools)
library(caret)
```

## Data Extraction

In this section, I scraped statcast data from Baseball Savant. For some reason, there is a cap of 40,000 rows one can scrape at a time, so I had to manually scrape data week-by-week to stay under the cap. After scraping the week-by-week data, I combined all of those data frames by year, then combined all of those yearly data frames to make one very large data frame called all_data_17_to_20.

Please bare with me as I am not the best at creating variable names. I also commented out the scraping code to improve the run time on this Shiny app.

```{r, eval = FALSE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
apr_data_2017_week1 <- scrape_statcast_savant_batter_all(start_date = "2017-04-02", end_date = "2017-04-08")
apr_data_2017_week2 <- scrape_statcast_savant_batter_all(start_date = "2017-04-09", end_date = "2017-04-15")
apr_data_2017_week3 <- scrape_statcast_savant_batter_all(start_date = "2017-04-16", end_date = "2017-04-22")
apr_data_2017_week4 <- scrape_statcast_savant_batter_all(start_date = "2017-04-23", end_date = "2017-04-29")
apr_data_2017_week4plus <- scrape_statcast_savant_batter_all(start_date = "2017-04-30", end_date = "2017-04-30")

all_data_2017 <- rbind(apr_data_2017_week1, apr_data_2017_week2, apr_data_2017_week3, apr_data_2017_week4, apr_data_2017_week4plus)
all_data_17_to_20 <- rbind(all_data_2017, all_data_2018, all_data_2019, all_data_2020)
```

After saving the data to a CSV, we'll read in the data and check it out.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
all_data_17_to_20 <- read_csv("all_events_data_17_to_20.csv")
head(all_data_17_to_20)
```

After noticing the data included pitchers with PA's, I needed to figure out a way to omit those players from the data set. I found that you can scrape the ["expected_statistics"](https://baseballsavant.mlb.com/leaderboard/expected_statistics?type=pitcher) leaderboard page by player position! I then scraped all non-pitcher data and combined it with the all_data_17_to_20 data frame. I tried setting the position parameter to position != 1, but that did not work out. I also joined in plate appearances to our main data frame.

```{r, eval = FALSE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
# Scraping in baseball savant data from 2017 to get Plate Appearances and Primary Position
statcast_scraper_2017_catchers <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 2, player_type = "batter")%>%
  mutate(primary_position = 2)
statcast_scraper_2017_1b <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 3, player_type = "batter")%>%
  mutate(primary_position = 3)
statcast_scraper_2017_2b <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 4, player_type = "batter")%>%
  mutate(primary_position = 4)
statcast_scraper_2017_3b <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 5, player_type = "batter")%>%
  mutate(primary_position = 5)
statcast_scraper_2017_ss <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 6, player_type = "batter")%>%
  mutate(primary_position = 6)
statcast_scraper_2017_lf <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 7, player_type = "batter")%>%
  mutate(primary_position = 7)
statcast_scraper_2017_cf <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 8, player_type = "batter")%>%
  mutate(primary_position = 8)
statcast_scraper_2017_rf <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 9, player_type = "batter")%>%
  mutate(primary_position = 9)
statcast_scraper_2017_dh <- scrape_savant_leaderboards(leaderboard = "expected_statistics", year = 2017, min_pa = 1, position = 10, player_type = "batter")%>%
  mutate(primary_position = 10)
```

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
statcast_all_batters <- read_csv("statcast_all_batters.csv")
head(statcast_all_batters)
```

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
statcast_all_batters%>%
  mutate(player_name = paste(first_name, last_name, sep = " "))%>%
  rename(game_year = year, batter = player_id)%>%
  dplyr::select(player_name, batter, game_year, pa, primary_position)->statcast_join

all_data_17_to_20_join <- left_join(all_data_17_to_20, statcast_join, by = c("player_name", "batter", "game_year"))%>%
  mutate(primary_position = ifelse(is.na(primary_position), 1, primary_position))

all_data_17_to_20_join%>%
  dplyr::select(game_date, player_name, batter,
         pitcher, events, game_type,
         game_year, hit_distance_sc, launch_angle,
         type, game_year, barrel, primary_position, pa)->xHR_data
```

## Choosing our Predictors

In order to build a predictive model, it makes sense to see which predictors have the strongest correlation to our response variable and construct a model based on those predictors. From prior knowledge, barrels per batted ball and fly ball rate have a strong correlation with HR/FB. Let's see if we can find any other model-worthy predictors.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
xHR_data%>%
  dplyr::select(player_name, batter, game_year, events, game_type, primary_position)%>%
  filter(events == "home_run" & game_type == "R")%>%
  group_by(player_name, batter, game_year, primary_position) %>% tally() %>% rename(home_runs = n)->hr_per_season_df

xHR_data%>%
  filter(events != "null" & hit_distance_sc >= 150 & game_type == "R" & launch_angle >= 25 & launch_angle <= 50)%>%
  group_by(game_year, player_name, batter, primary_position)%>% tally()%>% rename(fly_balls = n)->fb_per_season_df

hr_per_fb_rate_join <- left_join(fb_per_season_df, hr_per_season_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(hr_per_fb_rate = home_runs/fly_balls)

xHR_data%>%
  dplyr::select(player_name, batter, game_year, game_type, barrel, primary_position)%>%
  filter(game_type == "R" & barrel == 1)%>%
  group_by(player_name, batter, game_year, primary_position) %>% tally() %>% rename(barrel_count = n) ->barrels_per_season_df

xHR_data%>%
  dplyr::select(player_name, batter, game_year, game_type, type, primary_position)%>%
  filter(game_type == "R" & type == "X")%>%
  group_by(player_name, batter, game_year, primary_position) %>% tally() %>% rename(bip = n) ->bip_per_season_df

brl_rate_join <- left_join(barrels_per_season_df, bip_per_season_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(brl_per_bip_rate = barrel_count/bip)

fly_ball_rate_join <- left_join(fb_per_season_df, bip_per_season_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(fly_ball_rate = fly_balls/bip)

xHR_data%>%
  dplyr::select(player_name, batter, game_year, game_type, hit_distance_sc, launch_angle, primary_position)%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50)%>%
  group_by(player_name, batter, game_year, primary_position)%>%
  summarize(avg_fly_ball_dist = mean(hit_distance_sc))->avg_fly_ball_dist_df
```

From the plot below and using the max and min function on hc_x, we’re able to see that the total distance of the outfield is 246 feet 2 feet in left field to 248 feet in right field. If we split the outfield in thirds, we’re able to calculate the total pulled and oppo fly balls for RHH and LHH.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
all_data_17_to_20%>%
  dplyr::select(hc_x, hc_y, type)%>%
  filter(type == "X")->hit_data

sample_data <- hit_data[sample(nrow(hit_data), 5000), ]

coord_plot <- ggplot(data = sample_data, aes(x = hc_x, y = (hc_y)*-1)) + geom_point() + ggtitle("X and Y Coords of 5,000 BIP Sample")

ggplotly(coord_plot)

hit_data <- na.omit(hit_data)

max(hit_data$hc_x)
min(hit_data$hc_x)
```

```{r}
all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50)%>%
  group_by(game_year)%>%
  tally()%>%
  rename(all_fb_league = n)->all_fb_league

all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50 & stand == "R" & hc_x <= 82)%>%
  group_by(game_year)%>%
  tally()%>%
  rename(RHH_pulled_fb_league = n)->RHH_pulled_fb_league

all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50 & stand == "L" & hc_x >= 164)%>%
  group_by(game_year)%>%
  tally()%>%
  rename(LHH_pulled_fb_league = n)->LHH_pulled_fb_league

pulled_and_oppo_join_league <- left_join(RHH_pulled_fb_league, LHH_pulled_fb_league, by = c("game_year"))%>%
  left_join(all_fb_league, by = c("game_year"))%>%
  replace(is.na(.), 0)%>%
  mutate(all_pulled_fb_league = RHH_pulled_fb_league + LHH_pulled_fb_league)%>%
  mutate(pulled_fb_rate_league = all_pulled_fb_league/all_fb_league)
```

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50 & stand == "R" & hc_x <= 82)%>%
  group_by(player_name, batter, game_year)%>%
  tally()%>%
  rename(RHH_pulled_fb = n)->RHH_pulled_fb

all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50 & stand == "R" & hc_x >= 164)%>%
  group_by(player_name, batter, game_year)%>%
  tally()%>%
  rename(RHH_oppo_fb = n)->RHH_oppo_fb

all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50 & stand == "L" & hc_x >= 164)%>%
  group_by(player_name, batter, game_year)%>%
  tally()%>%
  rename(LHH_pulled_fb = n)->LHH_pulled_fb

all_data_17_to_20%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50 & stand == "L" & hc_x <= 82)%>%
  group_by(player_name, batter, game_year)%>%
  tally()%>%
  rename(LHH_oppo_fb = n)->LHH_oppo_fb

RHH_pulled_and_oppo_join <- left_join(RHH_pulled_fb, RHH_oppo_fb, by = c("player_name", "batter", "game_year"))%>%
  left_join(fb_per_season_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(RHH_pulled_fb_rate = (RHH_pulled_fb)/fly_balls)%>%
  rename(pulled_fb = RHH_pulled_fb,
         oppo_fb = RHH_oppo_fb,
         pulled_fb_rate = RHH_pulled_fb_rate)

LHH_pulled_and_oppo_join <- left_join(LHH_pulled_fb, LHH_oppo_fb, by = c("player_name", "batter", "game_year"))%>%
  left_join(fb_per_season_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(LHH_pulled_fb_rate = (LHH_pulled_fb )/fly_balls)%>%
  rename(pulled_fb = LHH_pulled_fb,
         oppo_fb = LHH_oppo_fb,
         pulled_fb_rate = LHH_pulled_fb_rate)

Both_pulled_and_oppo_df <- rbind(RHH_pulled_and_oppo_join, LHH_pulled_and_oppo_join)

Both_pulled_and_oppo_df%>%
  group_by(player_name, batter, game_year)%>%
  summarize(pulled_fb = sum(pulled_fb),
            oppo_fb = sum(oppo_fb))->Both_pulled_and_oppo_df

Both_fb_rate_join <- left_join(Both_pulled_and_oppo_df, fb_per_season_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(pulled_fb_rate = pulled_fb/fly_balls)%>%
  mutate(oppo_fb_rate = oppo_fb/fly_balls)

big_join <- left_join(hr_per_fb_rate_join, brl_rate_join, by = c("player_name", "batter", "game_year"))%>%
  left_join(fly_ball_rate_join, by = c("player_name", "batter", "game_year"))%>%
  left_join(Both_fb_rate_join, by = c("player_name", "batter", "game_year"))%>%
  left_join(avg_fly_ball_dist_df, by = c("player_name", "batter", "game_year"))%>%
  mutate(fly_ball_oppo_and_pulled_rate = pulled_fb_rate + oppo_fb_rate)%>%
  dplyr::select(game_year, player_name, batter, bip.x, fly_balls.x, hr_per_fb_rate, brl_per_bip_rate, fly_ball_rate, pulled_fb_rate, oppo_fb_rate, fly_ball_oppo_and_pulled_rate, avg_fly_ball_dist)%>%
  rename(bip = bip.x, fly_balls = fly_balls.x)

big_join <- na.omit(big_join)
```

```{r}
matrix <- cor(big_join[6:12])
cor_table <- round(matrix, 2)
cor_table
```

It looks like barrels per balls in play, average fly ball distance, pulled fly ball rate, and oppo and pulled fly ball rate correlate most strongly to HR/FB.

## Data Modeling

After comparing a few other linear regression models, here was the best model I constructed.

$HRperFlyBallrate = BRLperBIPrate + FlyBallrate + PulledFlyBallrate + AvgFlyBalldist$

## Model Validation

To reach this final model, I tested my model for multicollinearity and overfitting.

I began by splitting the 'big_join' data into training and testing datasets. The purpose of this is to use the training data to fit the linear regression model and the test data is used to evaluate the fit of the model. I split the data 70%/30% into training and testing sets.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
set.seed(101)
sample <- sample.split(big_join$player_name, SplitRatio = .7)
train <- subset(big_join, sample == TRUE)
test <- subset(big_join, sample == FALSE)

train%>%dplyr::select(game_year, player_name, batter, hr_per_fb_rate, brl_per_bip_rate, fly_ball_rate, pulled_fb_rate, avg_fly_ball_dist)->train

model1 <- lm(hr_per_fb_rate ~., data = train[4:8])
summary(model1)
```

Since I wasn't sure if there was overlap between my predictors, I tested for multicollinarity. By computing the variance inflation factor (or VIF), I measured how much variance of each regression coefficient is inflated due to multicollinearity in the model. As a rule of thumb, a VIF value that is between 5 to 10 indicates a problematic amount of collinearity. When faced with multicollinearity, one of the collinear variables should be removed, since the presense of multicollinearity implies that information that these variables provide about the response is redundant.

As you can see, there are no signs of multicollinearity with our predictors.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
car::vif(model1)
```

Overfitting is when a model is able to fit almost perfectly your training data, but is performing poorly on the new data.

I also performed the K-fold Cross-Validaiton technique that estimated the prediction and error and accuracy of our best model. The cross-validation algorithm divides our 'big_join' data into K subsets or (folds) of roughly equal size. Out of these K folds, one subset is used as the testing set while the other folds are used for training the model. The purpose of cross-validating is to eliminate any bias in our training data. I chose K = 10, so the cross-validation algorithm was repeated 10 times until the model is trained and tested on all subsets. The output of our K-fold cross-validation technique is that it generates overall prediction error by taking the average prediction errors in every iteration.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
train.control <- trainControl(method = "repeatedcv",
                              number = 10,
                              repeats = 3)

# Train the model
model1_test <- train(hr_per_fb_rate ~., data = train[4:8], method = "lm",
                trControl = train.control)

# Sumamrize the results
print(model1_test)
```

Here is our best linear regression model to predict xHR/FB.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
final_model <- model1_test$finalModel
```

Here are our predictions on our testing data.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
test$hr_per_fb_rate_estimations <- predict(final_model, test)
```

I also tested for overfitting. One way to test for overfitting is by comparing the predicted R-squared to the regular R-squared and observe if there is a sizeable difference. If there is a large discrepancy between the two values, then the model doesn't predict new observations as well as it fits the original data set. The results are not generalizable, and there's a good chance you're overfitting the model.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
model_R2 <- cor(model1$fitted.values, train$hr_per_fb_rate)^2
model_test_R2 <- cor(test$hr_per_fb_rate, test$hr_per_fb_rate_estimations)^2
Overfitting_value <- model_R2 - model_test_R2
Overfitting_value
```

As you can see there is little discrepancy between the predicted R-squared and regular R-squared indicating that there is no overfitting in our model.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
test <- left_join(test, statcast_join, by = c("game_year", "player_name", "batter"))%>%
  filter(game_year == 2020)%>%
  select(player_name, hr_per_fb_rate, hr_per_fb_rate_estimations, primary_position)

colors <- c("2" = "#FF6600", "3" = "#0066FF", "4" = "#0066FF", "5" = "#0066FF", "6" = "#0066FF",
            "7" = "#00FF00", "8" = "#00FF00", "9" = "#00FF00", "10" = "#000000")

p <- ggplot(test, aes(x = hr_per_fb_rate_estimations, y = hr_per_fb_rate, color = factor(primary_position))) + geom_point(aes(text=player_name)) + 
  labs(title = "Test Data (Small Sample Size): 2020 Real vs Estimated HR/FB Rates",
       x = "Expected HR/FB",
       y = "Actual HR/FB") + scale_color_manual(values = colors) + geom_abline()

ggplotly(p)
```

Great! Now that we have a strong xHR/FB model, we'll need to predict our predictors - Brl/BIP, Fly ball rate, Pulled fly ball rate, and average fly ball distance.

## Marcel Projections

I'm going to attempt to predict our predictive rates using the Marcel projection system. Following [this](https://www.beyondtheboxscore.com/2016/2/22/11079186/projections-marcel-pecota-zips-steamer-explained-guide-math-is-fun) example from beyondtheboxscore. The author, Henry Druschel, explains the origins of the Marcel projections and illustrated a concrete example which I followed to predict our predictors.

Here's a brief rundown of what the Marcel Projections system is:

It's a methodology created by Tom Tango back in 2004 that uses the last three years of player performance, with the most recent weighed more heavily to predict future performance. It factors in league-wide rates for whatever statistic you're attempting to predict. It regresses each player's past performance towards the mean by setting each player's PA to a rate of 1200. Marcel also adjusts projection by age. Unfortunately, for this exercise, I did not account for player ages, but could work on implementing this in the future. Although these calculations are basic, Marcel projections hold their own against the more robust and current projections systems.

As an walk-through, I'll project all 2021 barrel totals. We begin by multiplying each batter's 2018 barrels by 5, 2019 barrels by 4 and 2020 barrels by 3. Note, since 2020 was a shortened season, I scaled all player stats to the equivalence of 600 PA's.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
xBRL_2021_join <- left_join(barrels_per_season_df, statcast_join, by = c("player_name", "batter", "game_year"))

xBRL_2021_join%>%
  mutate(scaled_2020_barrel_count = ifelse(game_year == 2020, (600/pa)*barrel_count, barrel_count))%>%
  mutate(weighed_BRL = ifelse(game_year == 2020, scaled_2020_barrel_count*5,
                              ifelse(game_year == 2019, scaled_2020_barrel_count*4,
                                     ifelse(game_year == 2018, scaled_2020_barrel_count*3, scaled_2020_barrel_count))))->barrel_step_1
```

Next, for plate appearances, we multiply last year (2019) by 0.5 and two years ago (2018) by 0.1, and add both of them to 200.

```{r}
barrel_step_1%>%
  mutate(scaled_PA = ifelse(game_year == 2020, 200+(600*.5), ifelse(game_year == 2019, pa*.1, pa)))->barrel_step_2
```

Next, we calculate league-wide rates for barrels per plate appearance for each of the last 3 seasons (excluding pitchers), and multiply those by each player's plate appearances and the annual weights, to get the weighted mean of barrels for the average player with the average amount of playing time over the last three years.

```{r}
barrel_step_2%>%
  mutate(league_BRL_over_PA = ifelse(game_year == 2020, 0.050386431, ifelse(game_year == 2019, 0.049582878, ifelse(game_year == 2018, 0.045560363, 0))))->barrel_step_3
```

Then, we regress each player's past performance towards the mean, by setting it to a rate per 1,200 PA's taking the average of it and each player's performance, weighted by PA and calculating a per-plate appearance rate.

```{r}
barrel_step_3%>%
  mutate(scaled_league_PA = ifelse(game_year == 2020, 5*600*league_BRL_over_PA,
                                   ifelse(game_year == 2019, 4*pa*league_BRL_over_PA,
                                          ifelse(game_year == 2018, 3*pa*league_BRL_over_PA, pa*league_BRL_over_PA))))%>%
  mutate(weighed_PAs = ifelse(game_year == 2020, 600*5, ifelse(game_year == 2019, pa*4, ifelse(game_year == 2018, pa*3, pa))))->xBRL_2021_prep
```

Finally, we multiply that rate by each player's projected number of plate appearances. To maintain uniformity, all projections are based on 560 plate appearances which is the equivalence of scaling each player's PA's over the past 2 seasons = 560 PA's = (600)0.5 + (600)0.1 + 200.

```{r}
xBRL_2021_prep%>%
  group_by(player_name)%>%
  summarize(numerator1 = sum(scaled_league_PA[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            denominator1 = sum(weighed_PAs[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            scaled_to_1200 = ((numerator1/denominator1)*1200),
            sum_of_weighed_barrel_count = ((numerator1/denominator1)*1200) + sum(weighed_BRL[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            final_denom = 1200+denominator1,
            expected_BRL_per_pa = sum_of_weighed_barrel_count/final_denom,
            expected_BRL_2021 = expected_BRL_per_pa*sum(scaled_PA[game_year == 2020 | game_year == 2019]))->all_expected_barrel_count_2021

all_expected_barrel_count_2021 <- na.omit(all_expected_barrel_count_2021)%>%dplyr::select(player_name, expected_BRL_2021)
```

This process was repeated to calculate projected balls in play and fly balls. For average fly ball distance, I simply took the average fly ball distance of all players from the past 3 seasons.

```{r}
xHR_data%>%
  dplyr::select(player_name, batter, game_year, game_type, type)%>%
  filter(game_type == "R" & type == "X")%>%
  group_by(player_name, batter, game_year) %>% tally() %>% rename(bip = n) ->all_bip_17_to_20


xBIP_2021_join <- left_join(all_bip_17_to_20, statcast_join, by = c("player_name", "batter", "game_year"))%>%rename(bip_count = bip)

xBIP_2021_join%>%
  mutate(scaled_2020_bip_count = ifelse(game_year == 2020, (600/pa)*bip_count, bip_count))%>%
  mutate(weighed_BIP = ifelse(game_year == 2020, scaled_2020_bip_count*5,
                              ifelse(game_year == 2019, scaled_2020_bip_count*4,
                                     ifelse(game_year == 2018, scaled_2020_bip_count*3, scaled_2020_bip_count))))%>%
  mutate(scaled_PA = ifelse(game_year == 2020, 200+(600*.5), ifelse(game_year == 2019, pa*.1, pa)))%>%
  mutate(league_BIP_over_PA = ifelse(game_year == 2020, 0.661173428, ifelse(game_year == 2019, 0.674204894, ifelse(game_year == 2018, 0.682092914, 0))))%>%
  mutate(scaled_league_PA = ifelse(game_year == 2020, 5*600*league_BIP_over_PA,
                                   ifelse(game_year == 2019, 4*pa*league_BIP_over_PA,
                                          ifelse(game_year == 2018, 3*pa*league_BIP_over_PA, pa*league_BIP_over_PA))))%>%
  mutate(weighed_PAs = ifelse(game_year == 2020, 600*5, ifelse(game_year == 2019, pa*4, ifelse(game_year == 2018, pa*3, pa))))->xBIP_2021_prep

xBIP_2021_prep%>%
  group_by(player_name)%>%
  summarize(numerator1 = sum(scaled_league_PA[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            denominator1 = sum(weighed_PAs[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            scaled_to_1200 = ((numerator1/denominator1)*1200),
            sum_of_weighed_bip_count = ((numerator1/denominator1)*1200) + sum(weighed_BIP[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            final_denom = 1200+denominator1,
            expected_BIP_per_pa = sum_of_weighed_bip_count/final_denom,
            expected_BIP_2021 = expected_BIP_per_pa*sum(scaled_PA[game_year == 2020 | game_year == 2019]))->all_expected_bip_count_2021

all_expected_bip_count_2021 <- na.omit(all_expected_bip_count_2021)%>%dplyr::select(player_name, expected_BIP_2021)

xFB_2021_join <- left_join(fb_per_season_df, statcast_join, by = c("player_name", "batter", "game_year"))

xFB_2021_join%>%
  mutate(scaled_2020_fly_balls = ifelse(game_year == 2020, (600/pa)*fly_balls, fly_balls))%>%
  mutate(weighed_FB = ifelse(game_year == 2020, scaled_2020_fly_balls*5,
                             ifelse(game_year == 2019, scaled_2020_fly_balls*4,
                                    ifelse(game_year == 2018, scaled_2020_fly_balls*3, scaled_2020_fly_balls))))%>%
  mutate(scaled_PA = ifelse(game_year == 2020, 200+(600*.5), ifelse(game_year == 2019, pa*.1, pa)))%>%
  mutate(league_FB_over_PA = ifelse(game_year == 2020, 0.233557874, ifelse(game_year == 2019, 0.237421991, ifelse(game_year == 2018, 0.237443218, 0))))%>%
  mutate(scaled_league_PA = ifelse(game_year == 2020, 5*600*league_FB_over_PA,
                                   ifelse(game_year == 2019, 4*pa*league_FB_over_PA,
                                          ifelse(game_year == 2018, 3*pa*league_FB_over_PA, pa*league_FB_over_PA))))%>%
  mutate(weighed_PAs = ifelse(game_year == 2020, 600*5, ifelse(game_year == 2019, pa*4, ifelse(game_year == 2018, pa*3, pa))))->xFB_2021_prep

xFB_2021_prep%>%
  group_by(player_name)%>%
  summarize(numerator1 = sum(scaled_league_PA[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            denominator1 = sum(weighed_PAs[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            scaled_to_1200 = ((numerator1/denominator1)*1200),
            sum_of_weighed_fly_balls = ((numerator1/denominator1)*1200) + sum(weighed_FB[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            final_denom = 1200+denominator1,
            expected_fb_per_pa = sum_of_weighed_fly_balls/final_denom,
            expected_fb_2021 = expected_fb_per_pa*sum(scaled_PA[game_year == 2020 | game_year == 2019]))->all_expected_fly_balls_2021

all_expected_fly_balls_2021 <- na.omit(all_expected_fly_balls_2021)%>%dplyr::select(player_name, expected_fb_2021)

xpulledFB_2021_join <- left_join(Both_pulled_and_oppo_df, statcast_join, by = c("player_name", "batter", "game_year"))

xpulledFB_2021_join%>%
  mutate(scaled_2020_pulled_fly_balls = ifelse(game_year == 2020, (600/pa)*pulled_fb, pulled_fb))%>%
  mutate(weighed_pulled_FB = ifelse(game_year == 2020, scaled_2020_pulled_fly_balls*5,
                             ifelse(game_year == 2019, scaled_2020_pulled_fly_balls*4,
                                    ifelse(game_year == 2018, scaled_2020_pulled_fly_balls*3, scaled_2020_pulled_fly_balls))))%>%
  mutate(scaled_PA = ifelse(game_year == 2020, 200+(600*.5), ifelse(game_year == 2019, pa*.1, pa)))%>%
  mutate(league_pulled_FB_over_PA = ifelse(game_year == 2020, 0.2470532, ifelse(game_year == 2019, 0.2326054, ifelse(game_year == 2018, 0.2256887, 0))))%>%
  mutate(scaled_league_PA = ifelse(game_year == 2020, 5*600*league_pulled_FB_over_PA,
                                   ifelse(game_year == 2019, 4*pa*league_pulled_FB_over_PA,
                                          ifelse(game_year == 2018, 3*pa*league_pulled_FB_over_PA, pa*league_pulled_FB_over_PA))))%>%
  mutate(weighed_PAs = ifelse(game_year == 2020, 600*5, ifelse(game_year == 2019, pa*4, ifelse(game_year == 2018, pa*3, pa))))->xpulledFB_2021_prep

xpulledFB_2021_prep%>%
  group_by(player_name)%>%
  summarize(numerator1 = sum(scaled_league_PA[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            denominator1 = sum(weighed_PAs[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            scaled_to_1200 = ((numerator1/denominator1)*1200),
            sum_of_weighed_pulled_fly_balls = ((numerator1/denominator1)*1200) + sum(weighed_pulled_FB[game_year == 2020 | game_year == 2019 | game_year == 2018]),
            final_denom = 1200+denominator1,
            expected_pulled_fb_per_pa = sum_of_weighed_pulled_fly_balls/final_denom,
            expected_pulled_fb_2021 = expected_pulled_fb_per_pa*sum(scaled_PA[game_year == 2020 | game_year == 2019]))->all_expected_pulled_fly_balls_2021

all_expected_pulled_fly_balls_2021 <- na.omit(all_expected_pulled_fly_balls_2021)%>%dplyr::select(player_name, expected_pulled_fb_2021)

xHR_data%>%
  dplyr::select(player_name, batter, game_year, game_type, hit_distance_sc, launch_angle)%>%
  filter(game_type == "R" & hit_distance_sc >= 150 & launch_angle >= 25 & launch_angle <= 50)%>%
  group_by(player_name, batter)%>%
  summarize(avg_fly_ball_dist = mean(hit_distance_sc[game_year == 2018 |
                                                       game_year == 2019 |
                                                          game_year == 2020]))->avg_fly_ball_dist_df_for_2021

avg_fly_ball_dist_df_for_2021 <- na.omit(avg_fly_ball_dist_df_for_2021)

all_2021_projections <- left_join(all_expected_barrel_count_2021, all_expected_bip_count_2021, by = "player_name")%>%
  left_join(all_expected_fly_balls_2021, by = "player_name")%>%
  left_join(all_expected_pulled_fly_balls_2021, by = "player_name")%>%
  left_join(avg_fly_ball_dist_df_for_2021, by = "player_name")%>%
  mutate(expected_BRL_rate = expected_BRL_2021/expected_BIP_2021)%>%
  mutate(expected_FB_rate = expected_fb_2021/expected_BIP_2021)%>%
  mutate(expected_pulled_FB_rate = expected_pulled_fb_2021/expected_BIP_2021)%>%
  dplyr::select(player_name, batter, expected_fb_2021, expected_BRL_rate, expected_FB_rate, expected_pulled_FB_rate, avg_fly_ball_dist)

all_2021_projections <- na.omit(all_2021_projections)

all_2021_projections%>%
  rename(brl_per_bip_rate = expected_BRL_rate,
         fly_ball_rate = expected_FB_rate,
         pulled_fb_rate = expected_pulled_FB_rate,
         avg_fly_ball_dist = avg_fly_ball_dist)->all_2021_projections

all_2021_projections$est_hr_per_fb_2021 <- predict(model1, all_2021_projections[4:7])

all_2021_projections$expected_hr_2021 <- all_2021_projections$expected_fb_2021*all_2021_projections$est_hr_per_fb_2021
```

Here we have the 2021 expected home run leaderboard for all players with at least 1 PA in 2020.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
all_2021_projections%>%
  dplyr::select(player_name, batter, expected_hr_2021)->all_2021_projections

all_2021_projections%>%
  arrange(desc(expected_hr_2021))
```

And the more reasonable 2021 expected HR leaderboard for players with at least 50 PA's last season.

```{r, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
statcast_join%>%
  filter(game_year == 2020)->pas_2020

all_2021_projections_min_50_pa <- left_join(all_2021_projections, pas_2020, by = "player_name")

all_2021_projections_min_50_pa <- na.omit(all_2021_projections_min_50_pa)%>%
  filter(pa >= 50)%>%
  dplyr::select(player_name, batter.x, pa, expected_hr_2021)%>%
  rename(batter = batter.x)

all_2021_projections_min_50_pa%>%
  dplyr::select(player_name, batter, expected_hr_2021)%>%
  arrange(desc(expected_hr_2021))

# Full leaderboard can be found on my Github: proj_2021_HR_full_leaderboard.csv
```
